---
title: "506 HW1 1"
author: "Steven Leone"
format: html
editor: visual
---
This is the link to the repo: <https://github.com/leonestv-byte/Homework_1>.

This is the file for DATASCI 506 HW1.

```{r}
#| echo: false
# install.packages("readODS")
library(readODS)
library(dplyr)

```

## Exercise 1

a) Import the data into a data.frame in R. Use the dataset documentation to assign appropriate variable names. (Note: Downloading and unzipping the file can take place outside of your submitted document, but importing the file should be in the submission.)

```{r}
df <- read.table("data.csv", sep =",", header = TRUE)

descriptions <- read_ods("cultivars-description.ods")

```

b) Using the information provided in the dataset documentation: Verify that there are 320 observations in the dataset; 40 distinct cultivars; and all the observations were collected from two seasons. Produce a table showing the number of observations from each season.

```{r}
print(paste("The number of observations in the dataset is: ", nrow(df)))
```


```{r}
print(paste("The number of distinct cultivars is: ", nrow(descriptions)))
```

c) Use the data to answer the following questions:

What is the correlation between plant height (PH) and grain yield (GY)?

```{r}
cor(df$PH, df$GY)
```

Calculate the correlation between PH and GY separately for each season.

```{r}
season_1_cor <- cor(df[df$Season == 1, ]$PH, df[df$Season == 1, ]$GY,)
season_1_cor

```

```{r}
season_2_cor <- cor(df[df$Season == 2, ]$PH, df[df$Season == 2, ]$GY,)
season_2_cor

```
Which season has the higher correlation?

```{r}
season_with_higher_correlation <-ifelse(abs(season_1_cor) > abs(season_2_cor), "Season 1", "Season 2")
print(season_with_higher_correlation)
```


What is the grain yield of the soybean observation with the greatest plant height?

```{r}
df[df$PH == max(df$PH), ]$GY
```

What percentage of observations have a thousand seed weight (MHG) greater than the overall average thousand seed weight?

```{r}
nrow(df[df$MHG > mean(df$MHG), ]) / nrow(df)

```


Which cultivar has the highest average grain yield?

```{r}
result <- df |>
  group_by(Cultivar) |>       
  summarize(
    avg_value = mean(GY)     
  )

result[result$avg_value == max(result$avg_value), ]$Cultivar
```

d) Create a table showing the average value of each quantitative variable:

One row containing overall averages.

```{r}
averages_table <- data.frame(
  Repetition = mean(df$Repetition, na.rm = TRUE),
  IFP        = mean(df$IFP, na.rm = TRUE),
  PH         = mean(df$PH, na.rm = TRUE),
  
NLP = mean(df$NLP, na.rm = TRUE),

NGP = mean(df$NGP, na.rm = TRUE),

NGL = mean(df$NGL, na.rm = TRUE),

NS = mean(df$NS, na.rm = TRUE),

MHG = mean(df$MHG, na.rm = TRUE),

GY = mean(df$GY, na.rm = TRUE)

)



averages_table
```


One row for each season containing season-specific averages.

```{r}
df_season_1 = df[df$Season == 1,]
df_season_2 = df[df$Season == 2,]

averages_table_one <- data.frame(
  Repetition = mean(df_season_1$Repetition, na.rm = TRUE),
  IFP        = mean(df_season_1$IFP, na.rm = TRUE),
  PH         = mean(df_season_1$PH, na.rm = TRUE),
  
NLP = mean(df_season_1$NLP, na.rm = TRUE),

NGP = mean(df_season_1$NGP, na.rm = TRUE),

NGL = mean(df_season_1$NGL, na.rm = TRUE),

NS = mean(df_season_1$NS, na.rm = TRUE),

MHG = mean(df_season_1$MHG, na.rm = TRUE),

GY = mean(df_season_1$GY, na.rm = TRUE)

)



averages_table_one




averages_table_two <- data.frame(
  Repetition = mean(df_season_2$Repetition, na.rm = TRUE),
  IFP        = mean(df_season_2$IFP, na.rm = TRUE),
  PH         = mean(df_season_2$PH, na.rm = TRUE),
  
NLP = mean(df_season_2$NLP, na.rm = TRUE),

NGP = mean(df_season_2$NGP, na.rm = TRUE),

NGL = mean(df_season_2$NGL, na.rm = TRUE),

NS = mean(df_season_2$NS, na.rm = TRUE),

MHG = mean(df_season_2$MHG, na.rm = TRUE),

GY = mean(df_season_2$GY, na.rm = TRUE)

)



averages_table_two
```

e) Carry out a series of t-tests to examine whether the ngrain yield (GY) differs between the two seasons. State the null and the alternative hypotheses. Present the R output and interpret the results.

alternative hypothesis: alternative hypothesis: true difference in means between group 1 and group 2 is not equal to 0

null hypothesis: alternative hypothesis: true difference in means between group 1 and group 2 is equal to 0

```{r}
t_test_result <- t.test(GY ~ Season, data = df)

print(t_test_result)
```

We fail to reject the null. There is no statistically significant difference.

## Exercise 2


## Exercise 3

